How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-30 tse-paper/content.tex
add todays review and todos for me

\section{Systematic Literature Mapping Survey}
\label{sec:survey}

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
A field related to build log analysis is the analysis of system logs.
These system logs are produced during the runtime of a system, while
build logs are produced during the runtime of the build of the system.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} depicts two
exemplary lines from such a system log.

@@

\subsection{Literature Selection}
The first step in a systematic literature survey is to
establish which works one wants to cover and how to find
them~\cite{kitchenham2009systematic}.
Since the field of build log analysis is relatively young, we
surveyed as broad a population of scientific material as possible.
We did not limit the sources to specific Software Engineering
venues~\cite{petersen2015guidelines}, but included all sources
monitored by Google Scholar, including Bachelor's theses, Master's
theses, and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases because it is the search engine for scholarly material
with
the broadest and most current index, including preprints.

@@

% TODO Caro: check article vs.
work occurrences, balance use
Third, since we had been careful to include rather than exclude borderline
works, a deeper evaluation excluded another 15 articles from the 87
works left after the second filter phase.
Multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the analyzing of build logs.
This excluded another 11 articles.
In sum, we classified 61 articles in detail.

@@

\subsection{Results}
\addtolength{\tabcolsep}{-5pt}

@@

Of the 61 articles we inspected, 43 mentioned how they analyzed
build logs.
% TODO Caro: check method vs.
technique
However, only 16 of these described their method in detail.
We summarize the techniques that the articles
described for analyzing build logs in \Cref{tab:litsur:techniques}.

@@

\subsubsection{Which kinds of build logs are subject to
  analysis? (RQ 1.3)}
\Cref{fig:litsur:log_producer} shows the distribution of supported
build tools.
Of the articles we surveyed, 30\% analyzed
Maven~\cite{maven2019website},
16\% analyzed Gradle~\cite{gradle2020website},
and 10\% analyzed Ant build logs~\cite{ant2020website}.

A large amount of the articles analyze build logs
from a small number of build tools.
In most cases, this is because the format of build logs changes
~\cite{staahl2014modeling} and
therefore \textbf{``parsers must be specialized to each build and test
framework''}~\cite{tomassi2019bugswarm}.
Sometimes, aspects separate from the build log analysis motivate
the limitation to a specific build tool.
For example, Shi et al.
focussed on Maven
build logs because they also chose the PIT tool to calculate coverage
and mutation score~\cite{shi2018evaluating}.
PIT is only available as a Maven plugin.

Only 7\% of the proposed methods claimed to be language-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) used logs from a source that no
second article
targeted.
In 38\% of cases, the articles collected build logs from Travis
CI~\cite{travisci2019webpage};
18\% of the articles processed logs from TravisTorrent; and 8\% included
build logs from industrial projects.
The number of logs analyzed varied greatly between the articles
and we were not always able to
extract it confidently.
23 of the articles analyzed more than a thousand build logs, while one
investigated up to 122 million logs.
Of the surveyed articles, 31\% claim to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which
Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying},
and several other articles studied.

@@

\subsection{Discussion}
\label{sec:lit-sur:discussion}

% ----- descriptive / general -----
Our literature survey shows that despite its relatively young age
(first mentioned in a paper in 2001, vast majority after
2016), \textbf{build log analysis is an established technique}
in the literature, with 61 articles making use of it.

@@

% ----- information RQ 1.1 -----
In regards to the information which was targeted, \textbf{most build log
analysis approaches aim to retrieve the reason a build failed} by
targeting ``errors'' or ``failing tests''.
% which is what we also do in LogChunks
\textbf{The majority of approaches retrieves specific chunks of
text from the logs.}

% ----- technique RQ 1.2 -----
\textbf{Most log analysis is automated.} Automated approaches are
necessary as many studies target a large number of builds.
Articles
which explicitly mention a manual analysis of the logs, mostly
restrict their work to few logs or take a ``representative sample to
make manual analysis feasible''~\cite{zolfagharinia2017not}.
We observed \textbf{very little reuse of build log analysis tools},
even though
\textbf{developing automated analysis tools requires a lot of effort}.
Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone,'' while other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.
% not sure if we can state that (no data on this, more 'impression')
In addition, we rarely saw articles justifying why they use a certain
technique to analyze build logs as there is a general \emph{lack of
guidance on when to use which technique}.

% ----- kinds of build logs RQ1.3 -----
Several articles noted the variety in log formats for different build
tools, which requires customizing build log techniques for
every supported build tool.
% The following two sentences are passive, however flipped around
% have a different meaning (we want build tool -> study count)
\Cref{fig:litsur:log_producer} shows that only a few distinct
build tools are the target of
many studies and tools (roughly, area A).
The majority of build tools was however only targeted by one
article each (roughly, area B).
From this we follow that the large tail of
\textbf{seldomly-investigated
environments would benefit from the
existence of a more generic solution.}
In most cases, \textbf{researchers analyze build logs from open
source projects}, obtained through Travis CI.
In addition, we saw several articles that targeted the logs of only
one specific software project.

@@

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
An alternative to analyzing build logs, would be to integrate plugins
that report precise data about the build process, such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin} or
Deflaker by Bell et al.~\cite{bell2018deflaker}.
However, such plugins are not available for all build tools and they
do not enable the analysis of historical data---builds that ran before
the plugin was deployed cannot be analyzed.
In addition to that, when using such a plugin, one would have to know
beforehand which information should be reported from a build, whereas
build log analysis enables us to analyze the wide range of information
already baked into the build logs.
Their advantage is that they can precisely report the information which
is required.

@@

% Alt:
%\section{Analyzing Build Logs with Chunk Retrieval Techniques}
\section{Chunk Retrieval Techniques}
\label{sec:techniques}

In summary, the outcome of the literature survey show the need for
build log
analysis techniques that are \emph{automated},
\emph{agnostic of the build environment}
and can be \emph{configured with little effort}.

@@

We designed and implemented three chunk retrieval
techniques based on approaches
previously described by several articles:
PBE mimics regular expressions, while simplifying
their development.
We chose text similarity (CTS) as a common information retrieval
technique and implement the ad hoc approach of
searching for specific strings or keywords within KWS\@.

In this section, we define our notion of chunk
retrieval techniques and present each of the techniques
we developed in detail.
Following that, we illustrate how the techniques
operate on a concrete
example and sketch out their implementation.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}
In this section, we formalize our use of chunk retrieval techniques.
By this word, we refer to techniques that automatically
retrieve pieces of information that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.

Generally, chunk retrieval techniques
can also be used to classify build logs, e.g.\ by checking
which substring was retrieved.
Therefore, chunk retrieval techniques
cover the great majority of the types of log analysis we saw within
our survey.

@@

\emph{Running a
chunk retrieval technique} means executing a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.
\Cref{tab:techniques} summarizes and distinguishes the techniques
we present along the
characteristics described above.

\subsection{Program Synthesis by Example (PBE)}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide in an automatically
generated program.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
The \emph{PROSE library}~\cite{prose2019webpage} is the basis of
our implementation.
This library builds on the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable \emph{PROSE} to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the chunk of a build log that should be retrieved.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary is too complicated to be synthesized, PBE returns the
error message ``no program found.''

@@

\subsection{Common Text Similarity (CTS)}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
We investigate a chunk retrieval technique that
retrieves those lines from a build log which are most
similar to examples given by the user.

\subsubsection{Configuration}
To configure chunk retrieval through text similarity we chose to use
the same concept of examples as for PBE.
The lines of the output strings of the training examples define the
search query.
We represent each line as a term vector where each entry counts
how often a word appears within the line.
These vectors exist in a space ith as many dimensions as different
words appear in the log text.
This Vector Space Model~\cite{schutze2008introduction} defines
text lines as more similar if they contain the same words in the same
frequencies, meaning their term vectors point in a similar direction
in the vector space.
We use the cosine function to calculate the angle between two
term vectors.
The smaller the angle between term vectors, the more similar are the
corresponding lines.
This is called \emph{cosine similarity}~\cite{korenius2007principal}.

@@

\subsubsection{Application}
To retrieve the desired information from a build log, we parse the
whole text and process it in the same way as the search query.
The algorithm calculates the cosine
similarity~\cite{korenius2007principal} to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the outputs of the training examples determines how many of
the most similar lines our approach returns as the output of the retrieval
run.

\subsection{Keyword Search (KWS)}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason a build
failed while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.

@@

\subsubsection{Application}
For a retrieval run, we take a whole build log file as input and
search for all exact occurrences of the search keywords.
As keywords are
often not directly describing the desired information, but rather
appear close to the desired information, KWS also retrieves the lines
around the found keyword.
The number of surrounding lines retrieved is
the average of lines in the specification output of the training examples.

>>>>

\section{Systematic Literature Mapping Survey}
\label{sec:survey}
% tense: past

@@

\subsection{Distinction from System Log Analysis}
\label{sec:system-log-analysis}
% tense: present

A field related to build log analysis is the analysis of system logs.
These system logs are produced during the runtime of a system, while
build logs are produced during the runtime of the build of the system.
In terms of the contents of the logs, the main difference is that system
logs are fundamentally structured
through events.
Each line in a log file represents one event with a
set of fields: timestamp, verbosity level, and raw message
content~\cite{he2017towards}.
\Cref{lst:system-log} depicts two
exemplary lines from such a system log.

@@

\subsection{Literature Selection}
% tense: past

The first step in a systematic literature survey is to
establish which works one wants to cover and how to find
them~\cite{kitchenham2009systematic}.
Since the field of build log analysis is relatively young, we
surveyed as broad a population of scientific material as possible.
We did not limit the sources to specific Software Engineering
venues~\cite{petersen2015guidelines}, but included all sources
monitored by Google Scholar, including Bachelor's theses, Master's
theses, and academic slide decks.
We use Google Scholar as opposed to Scopus or
other databases because it is the search engine for scholarly material
with
the broadest and most current index, including preprints.

@@

% TODO Caro: check article vs work occurrences, balance use
Third, since we had been careful to include rather than exclude borderline
works, a deeper evaluation excluded another 15 articles from the 87
works left after the second filter phase.
Multiple articles were extensions of others.
Of these, we only considered the earliest article which reported on
the analyzing of build logs.
This excluded another 11 articles.
In sum, we classified 61 articles in detail.

@@

\subsection{Results}
% tense: past
\addtolength{\tabcolsep}{-5pt}

@@

Of the 61 articles we inspected, 43 mentioned how they analyzed
build logs.
% TODO Caro: check method vs technique
However, only 16 of these described their method in detail.
We summarize the techniques that the articles
described for analyzing build logs in \Cref{tab:litsur:techniques}.

@@

\subsubsection{Which kinds of build logs are subject to
  analysis? (RQ 1.3)}
As the structure of build logs can change greatly depending on the
build tools contributing to it~\cite{staahl2014modeling},
we were interested in which (variety) build tools are
investigate in literature.
\Cref{fig:litsur:log_producer} shows the distribution of
build tools whose logs were analyzed by the articles in our survey.
Of the articles, 30\% analyzed
Maven~\cite{maven2019website},
16\% analyzed Gradle~\cite{gradle2020website},
and 10\% analyzed Ant build logs~\cite{ant2020website}.

A large fraction of the articles analyze build logs
from a small number of build tools.
In most cases, this is because the format of build logs
% TODO Caro: check for tilde after space or newline
changes and
therefore ``parsers must be specialized to each build and test
framework''~\cite{tomassi2019bugswarm}.
Sometimes, aspects separate from the build log analysis motivate
the limitation to a specific build tool.
For example, Shi et al.
focussed on Maven
build logs because they also chose the PIT tool~\cite{coles2016pit}
to calculate coverage
and mutation score~\cite{shi2018evaluating}.
PIT is only available as a Maven plugin.
Only 7\% of the proposed methods claimed to be
build tool-agnostic, while
several described they are covering multiple source build tools.
The majority of articles (42\%) analyzed logs produced by a build tool
that no second article supported.

In 38\% of cases, the articles collected build logs from Travis
CI~\cite{travisci2019webpage};
18\% of the articles processed logs from TravisTorrent; and 8\% included
build logs from industrial projects.
The number of logs analyzed varied greatly between the articles
and we were not always able to
extract it confidently.

23 of the articles analyzed more than a thousand build logs, while one
investigated up to 122 million logs.
Of the surveyed articles, 31\% claim to have published their data to
enable further research.
However, we found no clear reuse of build log data sets with the
exception of TravisTorrent~\cite{beller2017travistorrent}, which
Vassallo et al.~\cite{vassallo2017a-tale},
Ghaleb et al.~\cite{ghaleb2019studying},
and several other articles studied.

@@

\subsection{Discussion}
\label{sec:lit-sur:discussion}
% tense: present

% TODO Caro: try emph instead of textbf
% ----- descriptive / general -----
Our literature survey shows that despite its relatively young age
(first mentioned in a paper in 2001, vast majority after
2016), \textbf{build log analysis is an established technique}
in the literature, with 61 articles making use of it.

@@

% ----- information RQ 1.1 -----
In regards to the information that is targeted, \textbf{most build log
analysis approaches aim to retrieve the reason a build failed} by
targeting ``errors'' or ``failing tests''.
% which is what we also do in LogChunks
\textbf{The majority of approaches retrieves specific chunks of
text from the logs.}

% ----- technique RQ 1.2 -----
\textbf{Most log analysis is automated.} Automated approaches are
necessary as many studies target a large number of builds.
Articles
that explicitly mention a manual analysis of the logs, mostly
restrict their work to few logs or take a ``representative sample to
make manual analysis feasible''~\cite{zolfagharinia2017not}.
We observed \textbf{very little reuse of build log analysis tools},
even though articles reported that
\textbf{developing automated analysis tools requires a lot of effort}.
Urli et al.~\cite{urli2018design} strongly discouraged from parsing
build logs for information as it is ``too error prone,'' while other
articles pointed to the effort of developing a custom tool.
Regular expressions are known to be tedious to
maintain~\cite{michael2019regexes}.
% not sure if we can state that (no data on this, more 'impression')
In addition, we rarely saw articles justifying why they use a certain
technique to analyze build logs as there is a general \textbf{lack of
guidance on when to use which technique}.

% ----- kinds of build logs RQ1.3 -----
Several articles noted the variety in log formats for different build
tools, which requires customizing build log techniques for
every supported build tool.
% The following two sentences are passive, however flipped around
% have a different meaning (we want build tool -> study count)
\Cref{fig:litsur:log_producer} shows that only a few distinct
build tools are the target of
many studies and tools (roughly, area A).
The majority of build tools was however only targeted by one
article each (roughly, area B).
From this we follow that the large tail of
\textbf{seldomly-investigated
environments would benefit from the
existence of a more generic solution.}
% In most cases, \textbf{researchers analyze build logs from open
% source projects}, obtained through Travis CI.
% In addition, we saw several articles that targeted the logs of only
% one specific software project.

@@

% TODO Moritz: is this okay (here)?
% maybe it's better to add it to Threats to Validity with the heading:
% \texbf{Modifying the Build Process.}
An alternative to analyzing build logs, would be to integrate plugins
that report precise data about the build process, such as the JUnit
plugin for Jenkins~\cite{jenkins2020junit-plugin} or
Deflaker by Bell et al.~\cite{bell2018deflaker}.
However, such plugins are not available for all build tools and they
do not enable the analysis of historical data---builds that ran before
the plugin was deployed cannot be analyzed.
In addition to that, when using such a plugin, one has to decide
beforehand which information should be reported from a build, whereas
build log analysis enables us to leverage the wide range of information
already baked into the build logs.
Their advantage is that they can precisely report the information which
is required.

@@

% Alt:
%\section{Analyzing Build Logs with Chunk Retrieval Techniques}
\section{Chunk Retrieval Techniques}
\label{sec:techniques}
% tense: past? except for general definitions / statements

In summary, the outcome of the literature survey shows the need for
build log
analysis techniques that are \emph{automated},
\emph{build tool-agnostic}
and can be \emph{configured with little effort}.

@@

We designed and implemented three chunk retrieval
techniques based on approaches
previously described by several articles:
\emph{Program Synthesis by Example (PBE)}
mimics regular expressions, while simplifying
their development.
From the field of information retrieval, we leverage a
\emph{Common Text Similarity (CTS)} approach
and implement the ad hoc approach of
searching for specific strings as \emph{Keyword Search (KWS)}.

In this section, we define our concept of chunk
retrieval techniques and present each of the techniques
we developed in detail.
Following that, we illustrate how the techniques
operate on a concrete
example and sketch out their implementation.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:crt-characteristics}
Before introducing each of the techniques, we formalize
we formalize our concept of chunk retrieval techniques.
By this phrase, we refer to techniques that automatically
retrieve pieces of information that appear literally in build
logs, i.e., techniques that do not aggregate, combine, or deduct
information.
We call such pieces of information in build logs
\emph{chunks}.

Generally, chunk retrieval techniques
can also be used to classify build logs, e.g.\ by checking
which substring was retrieved.
Therefore, they
cover the great majority of the types of log analysis we saw within
our survey.

@@

\emph{Running a
chunk retrieval technique} means applying a fully configured
technique to a build log in plain text format.
The technique then outputs a list that consists of substrings of the
build log text.
In \Cref{tab:techniques}, we summarize and distinguish the techniques
we present along the characteristics described above.

\subsection{Program Synthesis by Example (PBE)}
The concept of \emph{Programming by Example} aims to automatically
generate a program, by capturing the
intent of the user through examples that the user provides.
Leveraging this approach, we designed a chunk retrieval technique
for build logs.
The \emph{PROSE library}~\cite{prose2019webpage} is the basis of
our implementation.
This library builds on the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
% Both enable \emph{PROSE} to synthesize regular expression programs
% consistent~\cite{mitchell1982generalization} with a set of in-
% and output examples.
% The usage of example enables the user to configure a chunk
% retrieval without understanding the whole structure of
% the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the chunk of a build log that should be retrieved.
The PROSE program synthesis then tries to construct a program
consistent with all training
examples~\cite{mitchell1982generalization}.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary is too complicated to be synthesized, PBE returns the
error message ``no program found.''

@@

\subsection{Common Text Similarity (CTS)}
Text Similarity approaches are widely used to filter unstructured
textual software artifacts~\cite{runeson2007detection,
marcus2005recovery,antoniol2002recovering,mccarey2006recommending}.
These approaches fall into the broader area of information retrieval
techniques, which several of the articles from our literature survey
used to analyze build logs.
Similar to these approaches, we designed a chunk retrieval
technique that
retrieves those lines from a build log which are most
similar to examples given by the user.

\subsubsection{Configuration}
To configure CTS we chose to use
the same concept of examples as for PBE.
The lines of the targeted chunks from the training examples define the
search query.
We represent each line as a term vector where every entry counts
how often a word appears within the line.
These vectors exist in a space with as many dimensions as different
words appear in the log text.
This Vector Space Model~\cite{schutze2008introduction} defines
text lines as more similar if they contain the same words in the same
frequencies, making their term vectors point in a similar direction
in the vector space.
We use the cosine function to calculate the angle between two
term vectors.
The smaller the angle between term vectors, the more similar are the
corresponding lines.
This is called \emph{cosine similarity}~\cite{korenius2007principal}.

@@

\subsubsection{Application}
To retrieve the desired information from a build log, we parse the
whole text and process it in the same way as the search query.
The algorithm calculates the cosine
similarity to compare each line of the
build log with each line of the search query.
After summing up the
similarities of each build log line to all search query lines, we sort
the build log lines in decreasing similarity.
The average number of
lines in the chunks of the training examples determines how many of
the most similar lines our approach returns as the output of the retrieval
run.

\subsection{Keyword Search (KWS)}
When developers scavenge for a specific piece of information within a
large amount of unstructured information, a first ad hoc approach they
use is to search for related keywords.
Indeed, this was one of the
most common approaches we took when searching for the reason a build
failed while creating the \emph{LogChunks} data
set~\cite{brandt2020logchunks}.
To investigate the performance of this approach, we
designed a chunk retrieval technique based on keyword search.

@@

\subsubsection{Application}
For a retrieval run, we take a whole build log file as input and
search for all exact occurrences of the search keywords.
As keywords are
often not directly describing the desired information, but rather
appear close to the desired information, KWS also retrieves the lines
around the found keyword.
The number of surrounding lines retrieved is
the average number of lines in the chunks of the training examples.
